---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,results='hide',
                      fig.path = "man/readme/README-")
```

# REHAB - A R package to estimate predictive uncertainty by analysing REsidual Heteroscedasticity, Autocorrelation and Bias

## Introduction

The primary objective of `REHAB` if to estimate the total predictive uncertainty around a streamflow time series simulated by an hydrological model. This is achieved by analyzing the residuals (observed minus simulated values), and in particular:

1. their conditional bias, i.e. how the mean of residuals evolves as a function of the simulated streamflow (or some other predictors);
2. their heteroscedasticity, i.e. how the standard deviation of residuals evolves as a function of the simulated streamflow (or some other predictors);
3. their autocorrelation, i.e. how the lag-1 autocorrelation coefficient evolves as a function of the simulated streamflow gradient (or some other predictors).

## Installation

`REHAB` is not yet released on [CRAN](https://cran.r-project.org/), so the development version from [Github](https://github.com/BenRenard/REHAB) needs to be installed as follows:

```{r message=FALSE, eval=FALSE}
devtools::install_github('BenRenard/REHAB')
```

The package can then be loaded with:

```{r message=FALSE}
library(REHAB) 
```

## Basic Usage

To demonstrate the use of `REHAB`, we will use the `ArdecheRiver` dataset shipped with the package, which contains 10 years of observed and simulated daily streamflow as illustrated below. Simulated streamflows are from the [GR4J hydrological model](https://webgr.inrae.fr/eng/tools/hydrological-models/daily-hydrological-model-gr4j), available through the [airGR package](https://hydrogr.github.io/airGR/).

```{r Ardeche, fig.height=4, fig.width=8}
# Observed streamflow in black, simulated streamflow in red
plot(ArdecheRiver$date,ArdecheRiver$obs,xlab='Date',ylab='Streamflow [m3/s]',pch=19)
lines(ArdecheRiver$date,ArdecheRiver$sim,col='red')
```

The first step is to call the function `analyseResiduals`, which performs residual analysis and returns an object containing a whole bunch of information that will be detailed a bit later.

```{r res, fig.height=4, fig.width=8}
# Perform residual analysis on the first 5 years (1826 days)
res=analyseResiduals(ArdecheRiver$obs[1:1826],ArdecheRiver$sim[1:1826])
# Plotting the object will show an obs. vs. sim plot
plot(res)
```

To compute predictive uncertainty, the object `res` needs to be passed to the function `getUncertainty`.

```{r u, fig.height=4, fig.width=8}
# Compute uncertainty
u=getUncertainty(res)
# Plot it
plot(u,axisValues=ArdecheRiver$date[1:1826])
```

Note that by default, uncertainty is estimated around the simulated streamflow used for residual analysis in step one. It is possible to compute uncertainty around other simulated values, by simply passing them with the `sim=` argument.

```{r u2, fig.height=4, fig.width=8,warning=FALSE}
# Get uncertainty around simulated streamflow during the last 5 years
u=getUncertainty(res,sim=ArdecheRiver$sim[1827:3653])
# plot it
g=plot(u,axisValues=ArdecheRiver$date[1827:3653])
# Note that the returned plot is a ggplot2 object which can be further customized using ggplot2 utilities
g+ggplot2::xlim(as.Date('2003-10-01'),as.Date('2003-12-31'))+ggplot2::labs(y='Streamflow [m3/s]')
```

The object `u` returned by the `getUncertainty` is a list containing a bunch of information, in particular:

* `u$env` contains the uncertainty envelop and the predictive median

```{r env, results='markup'}
head(u$env)
```

* `u$spag` contains streamflow spaghettis, i.e. 1000 possible realizations of the streamflow time series, representing predictive uncertainty.

```{r spag, fig.height=4, fig.width=8}
# Plot 10 spaghettis during the first 60 days
matplot(t(u$spag[1:10,1:60]),type='l')
```

## Advanced Usage

Under the hood, `REHAB` performs polynomial regressions for the mean, the standard deviation and the lag-1 autocorrelation of residuals. The example below shows a few useful options to manipulate these regressions.

